---
title: "neopeptides_predictions"
author: "Dario Righelli"
date: "`r Sys.Date()`"
bibliography: references.bib  
output: 
    html_document:
        code_folding: hide
        toc: true
        toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Setting up libraries and working directory

```{r}
library(data.table)
library(dplyr)
library(tidyr)
library(readxl)
# setwd("~/neoantigens")
```



```{r}

neo <- read.table("../neoantigens_peptides_BigMHC_preds_07_02_2024.tsv", header=TRUE, sep="\t")

# pep_bmhc_el <- read.table("../neopep_bigMHC_EL.prd",
#                         header=TRUE,
#                         row.names = NULL, sep=",")
# pep_bmhc_im <- read.table("../neopep_bigMHC_IM.prd",
#                         header=TRUE,
#                         row.names = NULL, sep=",")

# neopep <- read.table("../neoantigens_peptides_BigMHC_preds_07_02_2024.tsv", header = TRUE, sep="\t")

pep_mmhc <- read_xlsx("../mergedData20240628_features.xlsx")
pep_mmhc <- as.data.frame(pep_mmhc)

```


```{r}
# pep_bmhc_el$mhc_pep <- paste0(pep_bmhc_el$mhc, "_", pep_bmhc_el$pep)
# pep_bmhc_im$mhc_pep <- paste0(pep_bmhc_im$mhc, "_", pep_bmhc_im$pep)
# head(pep_bmhc_im)
# head(pep_bmhc_el)
# head(pep_mmhc)
# head(neo)

neo1 <- neo[order(neo$mhc_pep),]
# identical(neo1$mhc_pep, pep_mmhc$mhc_pep)
neo12 <- cbind(neo1, pep_mmhc[,which(!colnames(pep_mmhc) %in% colnames(neo1))])
# saveRDS(neo12, file="../neopeptides_BigMHC_Prime_MixMHC_pred_2024_07_09.RDS")
# write.table(neo12, file="../neopeptides_BigMHC_Prime_MixMHC_pred_2024_07_09.tsv", row.names = FALSE, quote=FALSE, sep="\t")
```

# Explorative analysis

## Gene Expression

```{r}
library(corrplot)
library(pheatmap)
m <- cbind(neo12$TCGA_Cancer_expression, neo12$rnaseq_TPM,
           neo12$Sample_Tissue_expression_GTEx, 
           neo12$GTEx_all_tissues_expression_mean)
rownames(m) <- neo12$mhc_pep
colnames(m) <- c("TCGA", "RNAseq",
           "GTEx_Tissue", 
           "GTEx_Tissues")
mna <- m[-which(is.na(m), arr.ind = TRUE)[,1],]
mnau <- unique(mna)
dim(mnau)
cr <- cor(mnau)
corrplot(cr, method = 'number')
pmnau <- t(scale(t(log1p(mnau)), center = TRUE, scale = FALSE))
pheatmap(pmnau)
```


## PrimeScore by tgt

```{r}
library(ggplot2)
library(ggpubr)
t_test <- t.test(PrimeScore ~ tgt, data = neo12)

# Extract the p-value
p_value <- t_test$p.value
ggplot(neo12, aes(x = as.factor(tgt), y = PrimeScore)) +
    geom_boxplot() +
    labs(x = "Target (tgt)", y = "PrimeScore", title = "Boxplot of PrimeScore by Target") +
    theme_minimal()+
    annotate("text", x = 1.5, y = max(neo12$PrimeScore), 
             label = paste("p-value =", signif(p_value, digits = 7)), size = 5)

```


### Prime score distribution

```{r}
g1 <- ggplot(neo12, aes(x=PrimeScore)) + geom_histogram(binwidth = 0.05, color="black", fill="white") +theme_bw() 
g2 <- ggplot(neo12, aes(x=PrimeScore)) + geom_histogram(aes(y=..density..), color="black", fill="white") + geom_density(alpha=.2, fill="#FF6666") +theme_bw() 
g3 <- ggplot(neo12, aes(x=PrimeScore, color=tgt)) + geom_histogram( fill="white", alpha=0.5, position="identity")+theme_bw() 
ggarrange(g1, g2, g3)

```

## BigMHC-IM - BigMHC-EL 

correspondence between scores

```{r}
neo12$tgt <- as.factor(neo12$tgt)
ggplot(neo12, aes(x=BigMHC_EL, y=BigMHC_IM, mhc_pep=mhc_pep, tgt=tgt)) + geom_abline(intercept = 0, col="grey")+ geom_point(aes(col=tgt), size=0.5) +theme_bw()
```

## BigMHC-IM

### BigMHC-IM by tgt

```{r}
t_test <- t.test(BigMHC_IM ~ tgt, data = neo12)

# Extract the p-value
p_value <- t_test$p.value
ggplot(neo12, aes(x = as.factor(tgt), y = BigMHC_IM)) +
    geom_boxplot() +
    labs(x = "Target (tgt)", y = "BigMHC_IM", title = "Boxplot of BigMHC_IM by Target") +
    theme_minimal()+
    annotate("text", x = 1.5, y = max(neo12$BigMHC_IM), 
             label = paste("p-value =", signif(p_value, digits = 7)), size = 5)
```

### BigMHC-IM score distribution

```{r}
g1 <- ggplot(neo12, aes(x=BigMHC_IM)) + geom_histogram(binwidth = 0.05, color="black", fill="white") +theme_bw() 
g2 <- ggplot(neo12, aes(x=BigMHC_IM)) + geom_histogram(aes(y=..density..), color="black", fill="white") + geom_density(alpha=.2, fill="#FF6666") +theme_bw() 
g3 <- ggplot(neo12, aes(x=BigMHC_IM, color=tgt)) + geom_histogram( fill="white", alpha=0.5, position="identity")+theme_bw() 
ggarrange(g1, g2, g3)

```

## BigMHC-EL

### BigMHC-EL by tgt

```{r}
t_test <- t.test(BigMHC_EL ~ tgt, data = neo12)

# Extract the p-value
p_value <- t_test$p.value
ggplot(neo12, aes(x = as.factor(tgt), y = BigMHC_EL)) +
    geom_boxplot() +
    labs(x = "Target (tgt)", y = "BigMHC_EL", title = "Boxplot of BigMHC_EL by Target") +
    theme_minimal()+
    annotate("text", x = 1.5, y = max(neo12$BigMHC_EL), 
             label = paste("p-value =", signif(p_value, digits = 7)), size = 5)
```


### BigMHC-EL score distribution

```{r}
g1 <- ggplot(neo12, aes(x=BigMHC_EL)) + geom_histogram(binwidth = 0.05, color="black", fill="white") +theme_bw() 
g2 <- ggplot(neo12, aes(x=BigMHC_EL)) + geom_histogram(aes(y=..density..), color="black", fill="white") + geom_density(alpha=.2, fill="#FF6666") +theme_bw() 
g3 <- ggplot(neo12, aes(x=BigMHC_EL, color=tgt)) + geom_histogram( fill="white", alpha=0.5, position="identity")+theme_bw() 
ggarrange(g1, g2, g3)
```

## MixMHC

### MixMHC by tgt

```{r}
t_test <- t.test(MixMHCScore ~ tgt, data = neo12)

# Extract the p-value
p_value <- t_test$p.value
ggplot(neo12, aes(x = as.factor(tgt), y = MixMHCScore)) +
    geom_boxplot() +
    labs(x = "Target (tgt)", y = "MixMHCScore", title = "Boxplot of MixMHCScore by Target") +
    theme_minimal()+
    annotate("text", x = 1.5, y = max(neo12$MixMHCScore), 
             label = paste("p-value =", signif(p_value, digits = 7)), size = 5)
```


### MixMHC score distribution

```{r}
g1 <- ggplot(neo12, aes(x=MixMHCScore)) + geom_histogram(binwidth = 0.05, color="black", fill="white") +theme_bw() 
g2 <- ggplot(neo12, aes(x=MixMHCScore)) + geom_histogram(aes(y=..density..), color="black", fill="white") + geom_density(alpha=.2, fill="#FF6666") +theme_bw() 
g3 <- ggplot(neo12, aes(x=MixMHCScore, color=tgt)) + geom_histogram( fill="white", alpha=0.5, position="identity")+theme_bw() 
ggarrange(g1, g2, g3)
```


## Other Features

### Other features by target

```{r}
features <- colnames(neo12)[c(26:dim(neo12)[2])][-which(colnames(neo12)[c(26:dim(neo12)[2])] %in% c("BigMHC_EL", "BigMHC_IM", "PrimeScore", "MixMHCScore"))]

for (f in features){
  t_test <- t.test(neo12[[f]] ~ tgt, data = neo12)
  
  # Extract the p-value
  p_value <- t_test$p.value
  
  p=ggplot(neo12, aes(x = as.factor(tgt), y = neo12[[f]])) +
    geom_boxplot() +
    labs(x = "Target (tgt)", y = f, title = paste0("Boxplot of ", f, " by Target - pval ",
                                                   signif(p_value, digits = 4))) +
    theme_minimal()
    #annotate("text", x = 1.5, y = max(pepTableb$PrimeScore), 
             #label = paste("p-value =", signif(p_value, digits = 7)), size = 5)
  print(p)
}

```


### Other features distribution

```{r}

for (f in features){
    g1 <- ggplot(neo12, aes(x=neo12[[f]])) + geom_histogram(binwidth = 0.05, color="black", fill="white")+
      labs(x = f, title = paste0("Histogram of ", f))+theme_bw()
    g2 <- ggplot(neo12, aes(x=neo12[[f]])) + geom_histogram(aes(y=..density..), color="black", fill="white") + geom_density(alpha=.2, fill="#FF6666")+ labs(x = f, title = paste0("Density of ", f)) +theme_bw() 
    print(ggarrange(g1, g2))
}
```













